Download data

### 2024-03-13 scripts/download_dependent_variables.Rmd
playing with burning raster
rerunninng modelling for missed az strip species

@@

BIEN is pretty finicky these days and is prone to crashing. There is a paper by Maitner with Enquist, Wilson, Merow etc. about the need for open source devs to be rewarded academically, I take that has a sign of Brian stepping away from it. We will use a for loop to download from BIEN. 
```{r Perform bien queries, eval = F}

@@

@@

ob <- lapply(target_taxa, idig_records)
idig <- bind_rows(ob) %>% 
  filter(datecollected > '1950-01-01')

write.csv(idig, '../data/raw/occurrence/IDIG2.csv')
```

@@

target_symbols <- c('ADFA', 'ADFAF', 'ADFAO', 'BREL', 'BRELE', 'KRGR', 'LOHU2')
targets <- filter(targets, symbol %in% target_symbols) %>% 
  distinct()

@@

st_write(spp_richness, '../data/raw/occurrence/AIM/AIM_records2.shp', append = F)

@@

@@

>>>>

## these species missed, fixed on March 13th
target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, target_taxa)) %>% 
  pull(taxon)

targets <- target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')))

@@

BIEN is pretty finicky these days and is prone to crashing. There is a paper by Maitner with Enquist, Wilson, Merow etc. about the need for open source devs to be rewarded academically, I take that as a sign of Brian stepping away from it. We will use a for loop to download from BIEN.

```{r Perform bien queries, eval = F}

@@

```{r stitch together BIEN queries 3, eval = F}

p <- '../data/raw/occurrence/bien'
files <- file.path(p, list.files(p))
files <- files[grep(paste(paste0(gsub(' ', '_', target_taxa), '.csv'), collapse = '|'), basename(files))]

write.csv(bien_data, '../data/raw/occurrence/BIEN_records3.csv', row.names = FALSE)
```

@@

ob <- lapply(target_taxa, idig_records)
idig <- bind_rows(ob) %>% 
  filter(year > 1950)

write.csv(idig, '../data/raw/occurrence/IDIG3.csv')
```

@@

@@

spp_richness <- spp_richness[!st_is_empty(spp_richness),,drop=FALSE]

st_write(spp_richness, '../data/raw/occurrence/AIM/AIM_records3.shp', append = F)

@@

```{r Create consensus presence records 3, eval = F}

bien <- read.csv(file.path(base_p, 'BIEN_records3.csv')) %>% 
  st_as_sf(coords = c('longitude', 'latitude'), crs = 4326) %>% 
  mutate(date = as.Date(date_collected)) %>% 
  select(taxon = scrubbed_species_binomial, date)

idig <- read.csv(file.path(base_p, 'IDIG3.csv')) %>% 
  st_as_sf(coords = c('geopoint.lon', 'geopoint.lat'), crs = 4326) %>% 
  mutate(date = as.Date(datecollected)) %>% 
  select(taxon = scientificname, date)

aim <- st_read(file.path(base_p, 'AIM', 'AIM_records3.shp'), quiet = TRUE) %>% 
  mutate(date = as.Date(DatVstd)) %>% 
  select(taxon, date, PrmryKy) %>% 
  st_transform(4326)

gbif3 <- read.delim(file.path(base_p, 'gbif4', 'occurrence.txt'), na.strings = "") %>% 
  drop_na(decimalLongitude, decimalLatitude) %>% 
  filter(str_detect(decimalLatitude, '[A-Z]|[a-z]', negate = T)) %>% 
  st_as_sf(coords = c('decimalLongitude', 'decimalLatitude'), crs = 4326) %>% 
  mutate(date = as.Date(eventDate), .before = geometry, 
         coordinateUncertaintyInMeters = as.numeric(coordinateUncertaintyInMeters)) %>% 
  filter(coordinateUncertaintyInMeters <= 250) %>% 
  select(taxon = species, date)

records <- bind_rows(aim, bien, gbif3,  idig) %>% 
  drop_na(taxon) %>% 
  mutate(taxon = str_to_sentence(taxon), 
         taxon = if_else(taxon == 'Wyethia sagittata', 'Balsamorhiza sagittata', taxon)) %>%
  arrange(taxon)

splicies <- split(records, records$taxon)
thinned <- lapply(splicies, ONEperCELL)
thinned <- bind_rows(thinned)
st_write(thinned, '../data/raw/occurrence/thinned/thinned_occurrences3.shp', append = F)
repeated_occurrences <- lapply(splicies, multiplesPERcell)
repeated_occurrences <- purrr::discard(repeated_occurrences, ~nrow(.) < 2)
repeated_occurrences <- bind_rows(repeated_occurrences)
st_write(repeated_occurrences, '../data/raw/occurrence/repeated/repeated_occurrences3.shp', append = F)

@@

```{r remove points outside analytical area3, eval = F}
library(terra)
domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27, 50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection

thinned <- st_read('../data/raw/occurrence/thinned/thinned_occurrences3.shp', quiet = TRUE)
thinned <- thinned[lengths(st_disjoint(thinned, domain)) ==  0, ]
st_write(thinned, '../data/raw/occurrence/thinned/thinned_occurrences3.shp', append = FALSE)

repeated_occurrences <- st_read('../data/raw/occurrence/repeated/repeated_occurrences3.shp', quiet = TRUE)
repeated_occurrences <- thinned[lengths(st_disjoint(repeated_occurrences, domain)) ==  0, ]
st_write(repeated_occurrences, '../data/raw/occurrence/repeated/repeated_occurrences3.shp', append = FALSE)
```

### 2024-03-13 scripts/generate_RF_models_copy.Rmd
more rfe subset logic
writing process
rerunninng modelling for missed az strip species

```{r read in response data and model}
p <- '../data/raw/occurrence/combined_records'
p2f <- file.path(p, list.files(p, pattern = '.shp$'))
train_dat <- lapply(p2f, sf::st_read, quiet = TRUE)
train_dat <- bind_rows(train_dat) %>% 
  st_transform(crs(layers))

@@

cores <- parallel::detectCores()
lapply(splicies[3:335], modeller)

@@

>>>>

```{r read in response data and model, eval = F}
p <- '../data/raw/occurrence/combined_records'
p2f <- file.path(p, list.files(p, pattern = '.shp$'))
train_dat <- lapply(p2f, sf::st_read, quiet = TRUE)
train_dat <- bind_rows(train_dat) %>% 
  st_transform(crs(layers))

@@

cores <- parallel::detectCores()
lapply(splicies, modeller)

@@

Model the missing AZ strip species
```{r re run on missing az strip species, eval = F}

targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>%

  ## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

target_taxa1 <- c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')

target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, target_taxa1)) %>% 
  pull(taxon)

p <- '../data/raw/occurrence/combined_records'
p2f <- file.path(p, list.files(p, pattern = '.shp$'))
p2f <- p2f[grep(paste(paste0(gsub(' ', '_', target_taxa), '.shp'), collapse = '|'), basename(p2f))]

train_dat <- lapply(p2f, sf::st_read, quiet = TRUE)
train_dat <- bind_rows(train_dat) %>% 
  st_transform(crs(layers))

cores <- parallel::detectCores()
lapply(splicies, modeller) 
```

### 2024-03-13 scripts/generate_absences.Rmd
ran all models
rerunninng modelling for missed az strip species

records <- st_read('../data/raw/occurrence/thinned/thinned_occurrences.shp', quiet = TRUE)
records <- records %>% 
  group_by(taxon) %>% 
  filter(n() >= 15)
records <- split(records, f = records$taxon)

@@

st_write(out, '../data/raw/occurrence/absences/absences1km.shp', append = FALSE)

@@

collinear_probs <- read_rds('../results/VIF_PA_drop.rda')
bio <- usdm::exclude(bio, collinear_probs)

absences <- st_read('../data/raw/occurrence/absences/absences.shp', quiet = TRUE)
presences <- st_read('../data/raw/occurrence/thinned/thinned_occurrences.shp', quiet = TRUE) %>% 
  mutate(Occurrence = 1, 
         PtType = if_else(is.na(PrmryKy), 'AIM Presence', 'Opportunistic Presence'), .before = geometry)

@@

set.seed(27)
cn <- names(layers)
splicies <- split(recs, f = recs$taxon)

lapply(splicies[41:335], lda_PA_dropper, path = '../results/lda2/', col_names = cn)

which(names(splicies) == 'Asclepias speciosa')

>>>>

records <- st_read('../data/raw/occurrence/thinned/thinned_occurrences3.shp', quiet = TRUE)
records <- records %>% 
  group_by(taxon) %>% 
  filter(n() >= 15)
records <- split(records, f = records$taxon)

@@

st_write(out, '../data/raw/occurrence/absences/absences1km3.shp', append = FALSE)

@@

collinear_probs <- read_rds('../results/VIF_PA_drop.rda')
layers <- usdm::exclude(layers, collinear_probs)

absences <- st_read('../data/raw/occurrence/absences/absences1km3.shp', quiet = TRUE)
presences <- st_read('../data/raw/occurrence/thinned/thinned_occurrences3.shp', quiet = TRUE) %>% 
  mutate(Occurrence = 1, 
         PtType = if_else(is.na(PrmryKy), 'AIM Presence', 'Opportunistic Presence'), .before = geometry)

@@

recs <- recs %>% 
  mutate(
  taxon = str_replace(taxon, 'Aslcepias subulata', 'Asclepias subulata'),
  taxon = str_replace(taxon, 'Elymus violaceus', 'Elymus trachycaulus'),
  taxon = str_replace(taxon, 'Encelia actonii$', 'Encelia actoni'),
  taxon = str_replace(taxon, 'Hillaria', 'Hilaria')
  ) %>% 
  filter(! taxon %in% c('Ephedra antisyphilitica', 'Ephedra fasciculata', 'Ephedra coryi', 'Ephedra cutleri',
                        'Ephedra funerea', 'Ephedra nevadensis', 'Ephedra torreyana', 'Ephedra californica',
                        'Ephedra trifurca', 'Ephedra trifurcata', 'Ephedra viridis', 'Ephedra aspera'))

set.seed(27)
cn <- names(layers)
splicies <- split(recs, f = recs$taxon)
names(splicies)
lapply(splicies[8:23], lda_PA_dropper, path = '../results/lda3/', col_names = cn)

### 2024-03-13 scripts/subsetDataforCrews.Rmd
rerunninng modelling for missed az strip species

p2ts <- '/media/steppe/hdd/SeedPhenology/data/processed/timestamps'
ott1 <- lapply(
  X = teams_list[5], # 5 throws an error
  FUN = phen_tabulator,
  path = p2ts, 
  admu = admin_boundaries, 
  target_species = spp_needed
  ) # row 67

ott1[[1]]
```

>>>>

p2ts <- '/media/steppe/hdd/SeedPhenology/data/processed/timestamps'
ott1 <- lapply(
  X = teams_list, # 5 throws an error
  FUN = phen_tabulator,
  path = p2ts, 
  admu = admin_boundaries, 
  target_species = spp_needed
  ) # row 67

```
